---
title: "WAF model feature selection"
author: "Yevgeny V. Yorkhov"
date: "January 29, 2017"
output: html_document
---

This is log analysis of two different web sites. The goal of the analysis is (Internet bot detection)[https://en.wikipedia.org/wiki/Internet_bot]. There are two ideas explored in this document:

- Internet bots and people have different time interval between log records of the web site URL;
- Internet bots and people have difference in the resource type (html, jpg, css, etc...) they get access.

## Mean/Median and standard deviation (SD) analysis
```{r, echo=FALSE}
setwd('~eugen/work/waf_model')
```

Let's load data and have a quick look at it.
```{r, load_data}
load('data/bimicon_parsed.Rda')
df <- bimicon_data_parsed

# show data format
df[c(500,503,700),]

# Extract bot IP-addresses based on access to /robots.txt and word 
# 'bot' in the UserAgent string
bot_idx <- grep('bot', ignore.case = TRUE, df$ua)
bot <- df[bot_idx,]$ip
bot_idx <- grep('robot', df$url)
bot2 <- df[bot_idx,]$ip
bot <- unique(c(bot, bot2)); rm(bot2, bot_idx, bimicon_data_parsed)

df <- df[,c(1,2,3)]
Sys.setlocale("LC_TIME", "en_US.UTF-8")  # Modern Linux etc.
df$time <- strptime(df$time, "%Y-%m-%d %H:%M:%S") # convert time

```

Now we need to calculate time interval between log records for each IP-address.

```{r, calculate_time_difference}
################# Make TDIFF
### split dataframe by ip adresses
X <- split(df, df$ip)

### calculate time delays between clicks
Y <- X
TD <- lapply(Y, function(x) abs(c(0,diff(x$time, lag=1))))
X <- mapply(cbind, Y, tdiff = TD, SIMPLIFY = FALSE)

### merge data into a single dataframe
df <- X[[1]]
for (i in 2:length(X)) {
  df <- rbind(df, X[[i]])
  print(paste(i, length(X), sep = " "))
}
save(df, file='data/bimicon_tdiff.Rda')
```

## Plotting MEAN and SD
Having time interval between records we can extract features - mean and standard deviation. After extracting features we build a plot illustrating the difference between bots and people's access to web site. 

```{r, calculate_mean_and_sd_plot_it}
library(plyr)
df <- df[,c(1,4)]
l_IP <- split(df, df$ip)

### Calculate MEAN and SD
IP_ADDRS <- unlist(lapply(l_IP, function(x) as.character(x$ip[1])), use.names = FALSE)
COUNT <- unlist(lapply(l_IP, function(x) dim(x)[1]), use.names = FALSE)
MEDIAN <- unlist(lapply(l_IP, function(x) mean(x$tdiff)), use.names = FALSE)
SD <- unlist(lapply(l_IP, function(x) sd(x$tdiff)), use.names = FALSE)

ips <- as.data.frame(ips)
ip_click_counts <- as.data.frame(click_counts)

### Put it into a single dataframe
dml <- data.frame(ip=IP_ADDRS, ip_mean=MEDIAN, ip_sd=SD, count=COUNT)
dml <- dml[complete.cases(dml),]

# set label, 1 means - ip is a bot
dml$bot <- "man"
dml$bot[dml$ip %in% bot] <- "bot"

# remove outliers
dml <- dml[dml$ip_mean < 300,]
dml <- dml[dml$ip_sd < 300,]

library(ggplot2)
ggplot(dml, aes(x=ip_mean, y=ip_sd, colour=bot)) + geom_point(size=3)

```

## Plotting MEDIAN and SD
We may have a distribution with heavy tails so we a trying to use MEDIAN instead of MEAN as a feature.

```{r, calculate_median_and_sd_plot_it}
library(plyr)
df <- df[,c(1,3)]
l_IP <- split(df, df$ip)

### Calculate MEDIAN and SD
IP_ADDRS <- unlist(lapply(l_IP, function(x) as.character(x$ip[1])), use.names = FALSE)
COUNT <- unlist(lapply(l_IP, function(x) dim(x)[1]), use.names = FALSE)
MEDIAN <- unlist(lapply(l_IP, function(x) median(x$tdiff)), use.names = FALSE)
SD <- unlist(lapply(l_IP, function(x) sd(x$tdiff)), use.names = FALSE)

ips <- as.data.frame(ips)
ip_click_counts <- as.data.frame(click_counts)

### Put it into a single dataframe
dml <- data.frame(ip=IP_ADDRS, ip_mean=MEDIAN, ip_sd=SD, count=COUNT)
dml <- dml[complete.cases(dml),]

# set label, 1 means - ip is a bot
dml$bot <- "man"
dml$bot[dml$ip %in% bot] <- "bot"

# remove outliers
dml <- dml[dml$ip_mean < 300,]
dml <- dml[dml$ip_sd < 300,]

library(ggplot2)
ggplot(dml, aes(x=ip_mean, y=ip_sd, colour=bot)) + geom_point(size=2)
```


